Bioverse: Potentially Observable Exoplanet Biosignature Patterns Under the UV Threshold Hypothesis for the Origin of Life

### 2025-03-25 src/tex/ms.tex
discuss pre-main sequence effects; restructure section

All commonly investigated origins-of-life scenarios require water as a solvent;
we thus considered only rocky planets that may sustain liquid water on their surface, i.e., that occupy their host star's momentary \gls{HZ} during the above period, as well as at the epoch of observation.
To determine \gls{HZ} occupancy, we took into account the evolution of the host star's luminosity and \gls{HZ} boundaries.

The \gls{HZ} describes a region around a star where a planet with Earth's atmospheric composition and climate feedbacks can maintain liquid water on its surface~\citep[e.g.,][]{Ramirez2017,Ramirez2018,MolLous2022,Spinelli2023,Tuchow2023}.
Here, we adopted orbital distance estimates that define the \gls{HZ} as the region between the runaway greenhouse transition, where the stellar instellation cannot anymore be balanced through infrared cooling to space~\citep{Ingersoll1969}, and the maximum greenhouse limit, corresponding to the maximum distance at which surface temperatures allowing liquid water can be maintained through a CO$_2$ greenhouse effect~\citep{Kasting1991,Kasting1993,Underwood2003,Kopparapu2013,Kopparapu2014}.
\rev{The exact boundaries of the habitable zone are known to be sensitive to the star's luminosity, spectral type, the planet's mass, and the planet's atmospheric properties~\citep[e.g.,][]{Pierrehumbert2011,Ramirez2014,Ramirez2017,Ramirez2018b,Koll2019,Ramirez2018,Ramirez2020,Bonati2021,Chaverot2022,Turbet2023}.
Here, we adopted the commonly used parametrization in \citet{Kopparapu2014} to derive luminosity and planetary mass-dependent distance limits of the \gls{HZ} $a_\mathrm{inner}$ and $a_\mathrm{outer}$.
}

@@

To determine \gls{HZ} occupancy, we interpolated the stellar luminosity evolution grid of \citet{Baraffe1998} using a Clough Tocher interpolant~\citep[][see left panel of Figure~\ref{fig:hz_nuv_evo}]{Nielson1983,Alfeld1984} to compute the evolution of the inner (runaway greenhouse) and outer (maximum greenhouse) edges as a function of planet mass and stellar spectral type~\citep{Kopparapu2014}.
Being a local interpolation method, Clough Tocher enables rapid processing while producing a smooth interpolating surface that highlights local trends.
From this, we get each planet's epochs within and outside the \gls{HZ}.

>>>>

All commonly investigated origins-of-life scenarios require water as a solvent;
we thus considered only rocky planets that may sustain liquid water on their surface, i.e., that occupy their host star's momentary \gls{HZ} during the above period, as well as at the epoch of observation.
The \gls{HZ} describes a region around a star where a planet with Earth's atmospheric composition and climate feedbacks can maintain liquid water on its surface~\citep[e.g.,][]{Ramirez2017,Ramirez2018,MolLous2022,Spinelli2023,Tuchow2023}.
Here, we adopted orbital distance estimates that define the \gls{HZ} as the region between the runaway greenhouse transition, where the stellar instellation cannot anymore be balanced through infrared cooling to space~\citep{Ingersoll1969}, and the maximum greenhouse limit, corresponding to the maximum distance at which surface temperatures allowing liquid water can be maintained through a CO$_2$ greenhouse effect~\citep{Kasting1991,Kasting1993,Underwood2003,Kopparapu2013,Kopparapu2014}.
\rev{The exact boundaries of the habitable zone are known to be sensitive to the star's luminosity, spectral type, the planet's mass, and the planet's atmospheric properties~\citep[e.g.,][]{Pierrehumbert2011,Ramirez2014,Ramirez2017,Ramirez2018b,Koll2019,Ramirez2018,Ramirez2020,Bonati2021,Chaverot2022,Turbet2023}.
Here, we adopted the commonly used parametrization in \citet{Kopparapu2014} to derive luminosity and planetary mass-dependent distance limits of the \gls{HZ} $a_\mathrm{inner}$ and $a_\mathrm{outer}$.
}

@@

\revv{To determine \gls{HZ} occupancy, we took into account the evolution of the host star's luminosity and \gls{HZ} boundaries.
Specifically, we} interpolated the stellar luminosity evolution grid of \citet{Baraffe1998} using a Clough Tocher interpolant~\citep[][see left panel of Figure~\ref{fig:hz_nuv_evo}]{Nielson1983,Alfeld1984} to compute the evolution of the inner (runaway greenhouse) and outer (maximum greenhouse) edges as a function of planet mass and stellar spectral type~\citep{Kopparapu2014}.
Being a local interpolation method, Clough Tocher enables rapid processing while producing a smooth interpolating surface that highlights local trends.
From this, we get each planet's epochs within and outside the \gls{HZ}.

\revv{We note that in particular for late M dwarfs, the pre-main-sequence phase can reach into the age ranges considered here~\citep{Baraffe2015}. This may expose planets to elevated levels of radiation, potentially inducing runaway greenhouse effects, water loss, or atmospheric erosion~\citep{Ramirez2014,Luger2015}.
As such, the actual window of surface habitability may be shorter than the total time a planet spends in the habitable zone during its main-sequence lifetime.
We defer a detailed treatment of pre-main-sequence effects to future work.
}
